---
title: "HW 1"
author: Meghan Bellerose,
date: September 16, 2020
output: github_document
---

Solutions to P8105 homework 1.

```{r library, message = FALSE}
library(tidyverse)

```


## Problem 1

#### Create data frame

I created a data frame with the following elements: 

* A random sample of size 10 from a standard Normal distribution
* A logical vector indicating whether elements of the sample are greater than 0
* A character vector of length 10
* A factor vector of length 10 with 3 factor levels

```{r create_df}
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("low", "low", "low", "mod", "mod", "mod", "mod", "high", "high", "high"))
)

```

#### Take means

I took the mean of each variable in the data frame.

```{r mean_pull_r}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

We can take the mean of numbers and logical vectors, but not character or factor vectors.

#### Convert variables from one type to another

I converted the logical, character, and factor vectors to numeric.

```{r try_numeric}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```

When converted to numeric, the logical vector is reported as false = 0, true = 1, which explains why it generated a mean.

I then convert the logical vector in the following ways:

* To numeric and multiply the random sample by the result
* To a factor and multiply the random sample by the result
* To a factor and then convert the result to numeric and multiply the random sample by the result

```{r convert_types}
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp) * pull(prob1_df, samp)
```

## Problem 2

#### Penguins dataset and size scatterplot

I loaded the penguins dataset package.

```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

The penguins dataset has eight variables: `r names(penguins)`. The data frame has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

Data were collected on three species of penguins: `r levels(penguins$species)` living on the following three islands: `r levels(penguins$island)`. 

The penguin's mean bill length is `r round(mean(penguins$bill_length_mm, na.rm=TRUE), 1)` and mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm=TRUE), 1)` mm. The penguin's mean body mass is `r round(mean(penguins$body_mass_g, na.rm=TRUE), 1)` g.



Here is a scatter plot of penguin flipper length by bill length: 

```{r penguins_scatter_plot, warning=FALSE}
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm)) + geom_point(aes(color = species)) + labs(title = "Penguin bill and flipper length by species", x = "Bill length (mm)", y = "Flipper length (mm)", color = "Penguin species") 
```

Finally, I saved the penguins plot!

```{r save_plot, warning=FALSE}
ggsave("penguins_scatter_plot.pdf", height = 4, width = 6)
```
